Are Delayed Issues Harder to Resolve? Revisiting Cost-to-Fix of Defects throughout the Lifecycle

### 2016-09-12 phaseDelay4Emse.tex
Updates from ShareLaTeX

Like any good theory, DIE includes a rationale for why the expected results would be seen. McConnell mentions it as a ``common observation'' in the field and  summarizes the intuitive argument for why it should be so: 
\begin{quote}
''A small mistake in upstream work can affect large amounts of downstream work. A change to a single sentence in a requirements specification can imply changes in hundreds of lines of code spread across numerous classes or modules, dozens of test cases, and numerous pages of end-user documentation''~\cite{mcconnell01}. 
\end{quote}
Glass also endorses this rationale, asserting that ``requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' is ``really just common sense''~\cite{glass02}.  Other researchers
are just as adamant in asserting that that the delayed issue effect is a generally useful law of software engineering.
For example, what we call the delayed issued effect was listed at \#1 by Boehm and Basili in their ``Top 10 list'' of ``objective and quantitative data, relationships,
and predictive models that help
software developers avoid predictable pitfalls
and improve their ability to predict
and control efficient software projects''~\cite{boehm01}.

@@

\section{Motivation: ``DIE'' is commonly held, yet poorly documented}
\label{sect:belief}
% To assess the prevalence of DIE,   we conducted a survey of software engineers. If our surveyed practitioners make management decisions based on their
% understanding of SE theory, then the DIE  may well influence their decisions.
One reason that industrial practitioners and academics believe so strongly in the delayed issue effect is that it is often referenced
in the SE literature. 
%For example,
%we know that \fig{b81} has been presented to the graduate SE class at North Carolina State University, without quarrel or %critical comment, every year for the last decade.
Yet when we look at the literature, the evidence for
delayed issue effect is both very sparse and very old.

@@

Shull et al.~\cite{Shull02} conducted a literature survey and held a series of e-workshops with industry experts on fighting defects. Workshop participants from Toshiba and IBM reported cost-to-fix ratios between early lifecycle and post-delivery defects of 1:137 and 1:117 for large projects respectively~\cite{Shull02} -- but the raw data points were not provided and thus cannot be confirmed. A goal of agile methods 
is to reduce the difficulty associated with making changes later in the lifecycle~\cite{beck00}. Relatively little empirical data exists on this point.
Elssamadisy and Schalliol~\cite{Elssamadisy02} offer an anecdotal report on the growing, high cost of rework in a 50 person, three-year, 500KLOC Extreme Programming project as the project grew in size and complexity-- but again we cannot access their 
exact figures. This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access the data used to make prior conclusions.

@@

Some studies that report less-than large
 increase in the effort required to fix delayed issues.  Boehm~\cite{Boehm80} provides data suggesting that the cost-to-fix curve for small projects
 %, two student projects of 2000 deliverable source instructions, 
 is flatter than for large projects (the dashed line of Figure~\ref{fig:cost-to-fix}). Data from NASA's Johnson Space Flight Center, reported by Shull~\cite{Shull02}, found that the cost to fix certain non-critical classes of defects was fairly constant across lifecycle phases (1.2 hours on average early in the project, versus 1.5 hours late in the project). Royce~\cite{Royce98} studied  a million-line, safety-critical missile defense system. Design changes (including architecture changes) required approximately twice the effort of implementation and test changes, and the cost-to-fix in implementation and test phases increased slowly. Boehm~\cite{Boehm10} attributes this success to a development process focused on removing architecture risk early in the lifecycle. Willis et al. (\cite{willis1998hughes}, page 54) provide tables summarizing the effort to fix over 66,000 defects as a function of lifecycle phase injected and removed from multiple projects. The tables are partly obscured, but seem to provide the first large scale evidence that a)~DIE need not be exponential and b)~DIE need not be monotonically increasing.  Again, the data points from these studies are not available, and thus newer evidence both in favor of and contrary to the DIE cannot be evaluated.

@@

These projects took place at organizations in many countries and were conducted using  the Team Software Process (TSP$\textsuperscript{SM}$). Since 2000, the SEI has been teaching and coaching TSP teams. One of the authors (Nichols) has mentored software development teams and coaches around the world as they deploy TSP within their organizations since 2006.  The  most recent completions were in 2014.

@@

Organizations using TSP agree to provide their project data to the SEI for use in research. In return the SEI agrees that  data must not be traceable to its source. The data is collected at major project events: launch, interim checkpoints, and at project completion. The data from these TSP projects were collected and stored in the Software Engineering Measured Process Repository (SEMPR) at the SEI.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=2.5in]{defect_type_frequency.png}
\end{center}
\caption{Relative frequencies of these defect types seen in our TSP data.}\label{fig:dtypes}.
\end{figure}
In our TSP data,  the relative frequencies of these defect types are shown in \fig{dtypes}. Around a quarter of the fixes were simple documentation changes.
That said, 75\% of the changes are quote  elaborate;   e.g. fixes to function necessitates a careful reflection of the purpose of the code.

@@

One or more defects are reported against a single plan item in the time tracking logs, e.g., a review session, an inspection meeting, a test execution. The time-tracking logs against the plan items for a defect includes the time required to (a)~collect data and realize there is an error, 
(b)~prepare a fix,  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).
Since multiple defects can be recorded against a plan item, the time-to-fix a defect is defined as:
\[
\text{\emph{time-to-fix a defect}} := \frac{\text{time for defect plan item}}{\text{\# of defects in plan item}}
\]

@@

%\todo{should we add a citation to Petersen and a few more words on context.}
In this section we provide some descriptive statistics and discuss the of the projects from which this data was drawn and summarize some additional contextual information. The project contexts describe the conditions under which these measures were obtained, help determine relevance of the results, and may guide future data analysis with segmentation. Key attributes of the context include the business and application domains, product size, project duration,  work flows, team size, team management,  development and integration approaches, organization size, location or distribution, certifications, developer experience, programming languages and tools used.

We are unable at this time to provide all individual context data for each of the projects for several reasons. While the development data was recorded in tools and submitted in a structured form, context data was collected in less structured project questionnaires, site questionnaires, team member surveys, launch presentations and reports, post mortem presentations and reports. This data has not yet been mined from the submissions 1) because of the cost and effort required, 2) we are obligated to avoid providing any data that can identify projects (that is, the data must remain anonymous), and 3) the unstructured data is not always  complete  submitted.   Gathering more projects will make it easier to anonymize the data and overcome missing data problems. Interest in the data sets by the community may encourage our sponsor to fund additional data mining.  Nonetheless, much context is available from the project data  and we provide some additional context not included within the fact sheets.

@@

While the processes and work flows among these projects can vary, the logical order is described section \ref{development_phase} is followed. Development tasks such as requirements development, design, or code, are typically followed by an appraisal phase such as personal review or inspection. Effort and effectiveness of these activities vary among projects and developers.

@@

 \begin{figure}[ht]
\scriptsize
\centering
\begin{tabular}{lrrrrrrrl}
                                                                   & N   & Min & Q1   & Median & Q3    & Max   & Mean   &  Distribution \\\hline
Team size                                                          & 171 & 1   & 4    & 6      & 10    & 36    &  7.8    &  \includegraphics[width=15mm]{team_size.png} \\ 
Duration {[}days{]}                                                & 171 & 7   & 33   & 61     & 118   & 1918  & 107    & \includegraphics[width=15mm]{duration.png} \\
Added \& Modified LOC                                                 & 117 & 2   & 1125 & 4201   & 13092 & 88394 & 10259  & \includegraphics[width=15mm]{LOC.png} \\
Components                                                         & 171 & 8   & 26   & 49     & 107   & 4170  & 116.2  & \includegraphics[width=15mm]{components.png} \\
Year                                                               & 171 &     & 2011 & 2012   & 2013  & 2014  & 2011.9 & \includegraphics[width=15mm]{year.png}
\end{tabular}

@@

 Of the 171 projects in the sample, only 117 collected size data in lines of code, however all projects tracked effort and  the component counts with applied effort are provided. Other data is complete for all 171 projects. The projects were mostly of short duration and small to medium size. The median project began in 2012 lasted 61 days, produced 4200 Lines of Code, 49 components (modules or features). Duration ranged from 7 to 1918 days. Size ranged from minimal (this may represent a short maintenance project) to 88394. The earliest project was in 2006 and the most recent in 2014.

How many of these teams could be classified as "agile" is not clear because actual practices in the agile world can vary. We did not ask teams to self-identify, however we offer the following observations regarding characteristics commonly associated with agile behavior. 
\begin{itemize}
    \item all teams were self managed, defining work flows, practices, and schedules
    \item teams meet at least weekly to evaluate progress and replan
    \item most teams were small with a median size of 6 and a mean of 7.8. only 25\% of the teams were larger than 10 with a long tail on the distribution
    \item the median project lasted only 60 days, suggesting limited scope for each integration
\end{itemize}

@@

    The Scott-Knott ranker 
      was   recommended  by  Mittas and Angelis in a
    recent TSE'13 article~\cite{mittas13} and by Ghotra et al. in a recent
    ICSE'15 article~\cite{ghotra2015icse}
    Scott-Knott is a 
    top-down clustering approach used to rank different treatments. If that 
    clustering finds an ``interesting division'' of the data, then some 
    statistical test is applied to the two divisions to check if they are 
    statistically significant different. If so, Scott-Knott considers recurses into both 
    halves.
    Before Scott-Knot recurses,  however,  it apples some  statistical hypothesis test $H$ to check
    if $m,n$ are significantly different.

    To operationalize ``interesting'', Scott-Knott seeks the division of 
    of $l$ treatments into subsets of size $m,n$ of sizes $ls,ms,ns$ and median values
    $l.\mu, m.\mu, n.\mu$ (respectively) in order to maximize
 $\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.

   To operationalize $H$, we use both  bootstrap sampling  and the A12 effect size test.
   In other 
    words, we divide the data if \textit{both} bootstrap sampling and effect 
    size test agree that a division is statistically significant (with a 
    confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
    For a justification of the use of non-parametric bootstrapping, see Efron 
    \& Tibshirani~\cite[p220-223]{efron93}. For a justification of the use of 
    effect size tests see Shepperd and MacDonell~\cite{shepperd12a}; 
    Kampenes~\cite{kampenes07}; and Kocaguenli et 
    al.~\cite{Kocaguneli2013:ep}. These researchers warn that even if a 
    hypothesis test declares two populations to be ``significantly'' 
    different, then that result is misleading if the ``effect size'' is very 
    small. Hence, to assess the performance differences we first must rule out 
    small effects using A12, a test   recently endorsed by Arcuri and 
    Briand at ICSE'11~\cite{arcuri11}.

    To  apply Scott-Knott, we divided data into the phases $P_0$ where issues are introduced. Next, for each division, we separated all the issues that were removed at different subsequent issues $P_r \in \{P_1,P_2,..\}$.
    For each pair $P_0,P_r$, we build one treatment containing the  issue resolution times for     issues raises in $P_0$ and resolved in $P_r$. These treatments
    were then ranked by Scott-Knott.

@@

\multicolumn{7}{c}{~}  \\
1 &  Before   
      &   DesignInspect   & 10 &  14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   CodeReview      & 8 &  14 & 0.80 & \textcolor{black}{\rule{8mm}{2mm}}  \\
1 &   &   CodeInspect     & 10 &  16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   UnitTest        & 12 &   21 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &   &   IntTest         & 15 &   31 & 1.50 & \textcolor{black}{\rule{15mm}{2mm}}  \\
1 &   &   SysTest         & 11 &   22 & 1.10 & \textcolor{black}{\rule{12mm}{2mm}}  \\  
\hline\multicolumn{7}{c}{~}  \\
1 &  Planning     &   ReqtsReview     & 8 &  14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &               &   DesignInspect   & 11 &  13 & 1.38 & \textcolor{black}{\rule{13mm}{2mm}} \\
2 &               &   UnitTest        & 24 &   25 & 3.00 & \textcolor{black}{\rule{30mm}{2mm}} \\
\hline\multicolumn{7}{c}{~}  \\
1 &  Reqts   &   ReqtsReview     & 13 &  20 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} \\
1 &          &   ReqtsInspect    & 12 &   18 & 0.92 & \textcolor{black}{\rule{9mm}{2mm}}  \\
1 &          &   DesignReview    & 10 &   14 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}} \\
1 &          &   DesignInspect   & 9 &   15 & 0.69 & \textcolor{black}{\rule{6mm}{2mm}}   \\
1 &          &   CodeInspect     & 13 &  24 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} \\
1 &          &   UnitTest        & 10 &  17 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}}  \\
1 &          &   IntTest         & 33 &   42 & 2.54 & \textcolor{black}{\rule{25mm}{2mm}}  \\
1 &          &   SysTest         & 24 &  108 & 1.85 & \textcolor{black}{\rule{18mm}{2mm}}  \\
\hline\multicolumn{7}{c}{~}  \\
1 & Design   &   DesignReview    & 11 &   16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   DesignInspect   & 8 &   12 & 0.73 & \textcolor{black}{\rule{7mm}{2mm}}  \\
1 &   &   CodeReview      & 10 &   18 & 0.91 & \textcolor{black}{\rule{9mm}{2mm}}  \\
1 &   &   CodeInspect     & 9 &  14 & 0.82 & \textcolor{black}{\rule{8mm}{2mm}}  \\
1 &   &   UnitTest        & 11 &   18 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   IntTest         & 17 &   31 & 1.55 & \textcolor{black}{\rule{15mm}{2mm}}  \\
1 &   &   AcceptTest      & 14 &   19 & 1.27 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &   &   SysTest         & 13 &   18 & 1.18 & \textcolor{black}{\rule{12mm}{2mm}}  \\
\hline\multicolumn{7}{c}{~}  \\
1 &  Code   &   CodeReview    & 10 &  16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &    &   CodeInspect   & 10 &   15 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   UnitTest      & 12 &   20 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &    &   IntTest       & 14 &   25 & 1.40 & \textcolor{black}{\rule{14mm}{2mm}}  \\
1 &    &   AcceptTest    & 16 &   25 & 1.60 & \textcolor{black}{\rule{16mm}{2mm}}  \\
1 &   &   SysTest       & 13 &   20 & 1.30 & \textcolor{black}{\rule{13mm}{2mm}}  \\

@@

\end{center}
\caption{Median times to resolve issues seen in the SEI TSP data. The main effect here is that there is no clear exponential growth in the time to resolve issues as issues are injected in some phase $P_0$ then resolved in some subsequent issue $P_r$.
Shown where are the 50th percentiles of issue resolution times for issues injected in phase $P_O$ and resolved in phase $P_r$ (these values are calculated
by sorting all resolution time, then reporting the middle values of that sort)
The ``IQR'' column shows the ``inter-quartile range''; i.e. the range of values representing the 75th - 25th percentile range
The left-hand-side ``rank'' column shows the result of the Scott-Knott ranking procedure described in \tion{stats}. Note that
most treatments achieved the same ranks i.e. they were found to be insignificantly different  from each other (the one exception is the Planning:UnitTest results 
where UnitTests were ranked 2).
The right-hand-side bars
illustrate shows the releative sizes of the increases for the  50th (median)   percentile values. These increases are calculated with respect to the first value in each section ``Before, Planning, Reqts. Design, Code''.
These right-hand-side bars are unitless since they are ratios. For example,
on the last line, issues injected during coding and fixed in SysTest take 13 minutes (median) to resolve. This is 130\% more than the 10 minutes (median) required
to resolve coding issues during CodeReview. The right-hand-side bar visually represents that 130\%.}
\label{fig:raw}
\end{figure}

@@

\textit{Reliability of measures}: The base measures in this study described in \tion{data-collection} are \textit{defects} recorded in TSP defect logs and \textit{time} reported in time tracking logs. The primary threats to the reliability of these measures are: that the definition of a defect varies between projects and that time is not reported accurately or consistently. The reliability of the time reporting is discussed in \tion{data_integrity}. Time is reported on a level of minutes. We do not have an assessment of the error margin for time reporting. Some developers are less precise with time or estimates. Nonetheless defect time-to-fix must sum to less than the total effort in the phase and time log time stamps must be consistent with the time stamps and phase in the defect log. Based on time stamp consistency and conformance with Benford distributions, our experience suggests greater than 90\% of the time log data is recorded in real time. The fidelity and consistency of data will be subject of a future paper.

>>>>

Like any good theory, DIE includes a rationale for why the expected results would be seen. McConnell mentions it as a ``common observation'' in the field and  summarizes the intuitive argument for why it should be so: 
\begin{quote}
''A small mistake in upstream work can affect large amounts of downstream work. A change to a single sentence in a requirements specification can imply changes in hundreds of lines of code spread across numerous classes or modules, dozens of test cases, and numerous pages of end-user documentation''~\cite{mcconnell01}. 
\end{quote}
Glass also endorses this rationale, asserting that ``requirements errors are the most expensive to fix when found during production but the cheapest to fix early in development'' is ``really just common sense''~\cite{glass02}.  Other researchers
are just as adamant in asserting that the delayed issue effect is a generally useful law of software engineering.
For example, what we call the delayed issued effect was listed at \#1 by Boehm and Basili in their ``Top 10 list'' of ``objective and quantitative data, relationships,
and predictive models that help
software developers avoid predictable pitfalls
and improve their ability to predict
and control efficient software projects''~\cite{boehm01}.

@@

\section{Motivation: ``DIE'' is commonly held, yet poorly documented}
\label{sect:belief}
% To assess the prevalence of DIE,   we conducted a survey of software engineers. If our surveyed practitioners make management decisions based on their
% understanding of SE theory, then the DIE  may well influence their decisions.
One reason that industrial practitioners and academics believe so strongly in the delayed issue effect is that it is often referenced
in the SE literature. 
%For example,
%we know that \fig{b81} has been presented to the graduate SE class at North Carolina State University, without quarrel or %critical comment, every year for the last decade.
Yet when we look at the literature, the evidence for
delayed issue effect is both very sparse and very old.
As  shown in this section,
a goal of agile methods 
is to reduce the difficulty associated with making changes later in the lifecycle~\cite{beck00}. Yet, as shown below,
relatively little empirical data exists on this point.

@@

Shull et al.~\cite{Shull02} conducted a literature survey and held a series of e-workshops with industry experts on fighting defects. Workshop participants from Toshiba and IBM reported cost-to-fix ratios between early lifecycle and post-delivery defects of 1:137 and 1:117 for large projects respectively~\cite{Shull02} -- but the raw data points were not provided and thus cannot be confirmed. 
Elssamadisy and Schalliol~\cite{Elssamadisy02} offer an anecdotal report on the growing, high cost of rework in a 50 person, three-year, 500KLOC Extreme Programming project as the project grew in size and complexity-- but again we cannot access their 
exact figures. This was a common theme in the literature reviewed for this paper-- i.e.  that  it was no longer possible to access the data used to make prior conclusions.

@@

Some studies report less-than large
 increase in the effort required to fix delayed issues.  Boehm~\cite{Boehm80} provides data suggesting that the cost-to-fix curve for small projects
 %, two student projects of 2000 deliverable source instructions, 
 is flatter than for large projects (the dashed line of Figure~\ref{fig:cost-to-fix}). Data from NASA's Johnson Space Flight Center, reported by Shull~\cite{Shull02}, found that the cost to fix certain non-critical classes of defects was fairly constant across lifecycle phases (1.2 hours on average early in the project, versus 1.5 hours late in the project). Royce~\cite{Royce98} studied  a million-line, safety-critical missile defense system. Design changes (including architecture changes) required approximately twice the effort of implementation and test changes, and the cost-to-fix in implementation and test phases increased slowly. Boehm~\cite{Boehm10} attributes this success to a development process focused on removing architecture risk early in the lifecycle. Willis et al. (\cite{willis1998hughes}, page 54) provide tables summarizing the effort to fix over 66,000 defects as a function of lifecycle phase injected and removed from multiple projects. The tables are partly obscured, but seem to provide the first large scale evidence that a)~DIE need not be exponential and b)~DIE need not be monotonically increasing.  Again, the data points from these studies are not available, and thus newer evidence both in favor of and contrary to the DIE cannot be evaluated.

@@

These projects took place at organizations in many countries and were conducted using  the Team Software Process (TSP$\textsuperscript{SM}$). Since 2000, the SEI has been teaching and coaching TSP teams. One of the authors (Nichols) has mentored software development teams and coaches around the world as they deploy TSP within their organizations since 2006.  The  most recent completions were in 2015.

@@

Organizations using TSP agree to provide their project data to the SEI for use in research. In return the SEI agrees that  data must not be traceable to its source. The data are collected at major project events: launch, interim checkpoints, and at project completion. The data from these TSP projects were collected and stored in the Software Engineering Measured Process Repository (SEMPR) at the SEI.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=2.5in]{defect_type_frequency.png}
\end{center}
\caption{Relative frequencies of these defect types seen in our TSP data.}\label{fig:dtypes}.
\end{figure}
In our TSP data,  the relative frequencies of these defect types are shown in \fig{dtypes}. Around a quarter of the fixes were simple documentation changes.
That said, 75\% of the changes are quite elaborate;   e.g. fixes to function necessitates a careful reflection of the purpose of the code.

@@

In the TSP, defect data includes the affected artifact, the estimated developer fix effort (find and fix), the lifecycle phases in which the defect was injected and removed, and the developer who implemented the fix. In the database, the task is associated with a plan item. Defects (one or more) are recorded in the defect log and associated with the plan item (task) in the time tracking logs. For example, a review session, an inspection meeting, or a test would be plan items associated with some product component. When defects are found and fixed, the time recorded in the time-tracking logs against the plan items includes the direct effort time (stop watch rather than wall clock time) required to (a)~collect data and realize there is an error, 
(b)~prepare a fix,  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests). Although we have explicit estimates of "find and fix" effort for each defect, this fails to account for the full costs (e.g. meeting time or test execution). Because the vast majority of defects are removed in explicit removal phases, we chose to estimate defect cost using the entire time in removal phases divided by the number of defects. We recognize that this approach can exaggerate cost per defect for cases with few defects and large overhead effort, such large test suites or slow running tests that require continuous developer attention.  Nonetheless, this approach provides a better comparison between early removals from inspections later removals from test. The result will be a time per defect that is greater than the directly measured "find and fix" time, but smaller than the wall clock or calendar time.

Since multiple defects can be recorded against a plan item, the time-to-fix a defect is defined as:
\[
\text{\emph{time-to-fix a defect}} := \frac{\text{time for defect plan item}}{\text{\# of defects in plan item}}
\]

@@

%\todo{should we add a citation to Petersen and a few more words on context.}
In this section we provide some descriptive statistics, discuss the projects from which this data was drawn, summarize some additional contextual information. The project contexts describe the conditions under which these measures were obtained, help determine relevance of the results, and may guide future data analysis with segmentation. Key attributes of the context include the business and application domains, product size, project duration,  work flows, team size, team management,  development and integration approaches, organization size, location or distribution, certifications, developer experience, programming languages and tools used.

We are unable at this time to provide all individual context data for each of the projects for several reasons. While the development data was recorded in tools and submitted in a structured form, context data was collected in less structured project questionnaires, site questionnaires, team member surveys, launch presentations and reports, post mortem presentations and reports. This data has not yet been mined from the submissions 1) because of the cost and effort required, 2) we are obligated to avoid providing any data that can identify projects (that is, the data must remain anonymous), and 3) the unstructured data may not be complete when submitted.   Gathering more projects will make it easier to anonymize the data and overcome missing data problems. Interest in the data sets by the community may encourage our sponsor to fund additional data mining.  Nonetheless, much context is available from the project data  and we provide some additional context not included within the fact sheets.

@@

While the processes and work flows among these projects can vary, the logical order described in section \ref{development_phase} is followed. Development tasks such as requirements development, design, or code, are typically followed by an appraisal phase such as personal review or inspection. Effort and effectiveness of these activities vary among projects and developers.

@@

 \begin{figure}[ht]
\scriptsize
\centering
\renewcommand{\arraystretch}{1}%
\begin{tabular}{lrrrrrrrl}
                                                                   & N   & Min & Q1   & Median & Q3    & Max   & Mean   &  Distribution \\\hline
Team size                                                          & 171 & 1   & 4    & 6      & 10    & 36    &  7.8    &  \includegraphics[width=25mm]{team_size.png} \\ 
Duration {[}days{]}                                                & 171 & 7   & 33   & 61     & 118   & 1918  & 107    & \includegraphics[width=25mm]{duration.png} \\
Added \& Modified LOC                                                 & 117 & 2   & 1125 & 4201   & 13092 & 88394 & 10259  & \includegraphics[width=25mm]{LOC.png} \\
Components                                                         & 171 & 8   & 26   & 49     & 107   & 4170  & 116.2  & \includegraphics[width=25mm]{components.png} \\
Project Initiation Year                                                       & 171 &    2006 & 2011 & 2012   & 2013  & 2014  & 2011.9 & \includegraphics[width=25mm]{year.png}
\end{tabular}

@@

 Of the 171 projects in the sample, only 117 collected size data in lines of code, however all projects tracked effort and  the component counts with applied effort are provided. Other data are complete for all 171 projects. The projects were mostly of short duration and small to medium size. The median project began in 2012 lasted 61 days, produced 4,200 Lines of Code, 49 components (modules or features). Duration ranged from 7 to 1,918 days. Size ranged from minimal (this may represent a short maintenance project) to 88,394. The earliest project was in 2006 and the most recent in 2014.

How many of these teams could be classified as "agile" is not clear because actual practices in the agile world can vary. We did not ask teams to self-identify, however we offer the following observations regarding characteristics commonly associated with agile behavior. 
\begin{itemize}
    \item all teams were self managed, defining work flows, practices, and schedules
    \item teams met at least weekly to evaluate progress and re-plan
    \item most teams were small with a median size of 6 and a mean of 7.8; only 25\% of the teams were larger than 10 with a long tail on the distribution
    \item the median project lasted only 60 days, suggesting limited scope for each integration
\end{itemize}

@@

    Scott-Knott allows for a direct answer to the following questions: given an issue
    raised at phase $i$ and resolved at phase $\forall j,k \in \{i,i+1,i+2,...\}$, is it
    true that that the time to resolve issues at $j$ is significantly different 
    to the time to resolve issues at $k$? 
    \be
    \item Note that if $j,k$ times are significantly different, then we can compare
    the median values to say (e.g.) resolution time at phase $k$ is 3 times slower than phase $j$.
    \item Note also that if all times $j,k$ are {\em not} significantly different
    then we say that the phases all {\em rank} the same (and we denote this by setting
    all such {\em ranks} to 1).
    \ee
    In the following results, we nearly always encountered the second case;
    i.e.   the times to resolve issues at different times were usually {\em not significantly}
    different.

    As to technical details of the Scott-Knott methods, this ranker 
      was   recommended  by  Mittas and Angelis in a
    recent TSE'13 article~\cite{mittas13} and by Ghotra et al. in a recent
    ICSE'15 article~\cite{ghotra2015icse}
    Scott-Knott is a 
    top-down clustering approach used to rank different treatments. If that 
    clustering finds an ``interesting division'' of the data, then some 
    statistical test is applied to the two divisions to check if they are 
    statistically significant different. If so, Scott-Knott considers recurses into both 
    halves.
    Before Scott-Knot recurses,  however,  it applies some  statistical hypothesis test $H$ to check
    if $m,n$ are significantly different.

    To operationalize ``interesting'', Scott-Knott seeks the division of 
    $l$ treatments into subsets of size $m,n$ of sizes $ls,ms,ns$ and median values
    $l.\mu, m.\mu, n.\mu$ (respectively) in order to maximize
 $\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2$.

   To operationalize $H$, we use both  bootstrap sampling  and Vargha and Delaney's A12 test A12 effect size test.
   In other 
    words, we divide the data if \textit{both} bootstrap sampling and effect 
    size test agree that a division is statistically significant (with a 
    confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
    For a justification of the use of non-parametric bootstrapping, see Efron 
    \& Tibshirani~\cite[p220-223]{efron93}. For a justification of the use of 
    effect size tests see Shepperd and MacDonell~\cite{shepperd12a}; 
    Kampenes~\cite{kampenes07}; and Kocaguenli et 
    al.~\cite{Kocaguneli2013:ep}. These researchers warn that even if a 
    hypothesis test declares two populations to be ``significantly'' 
    different, then that result is misleading if the ``effect size'' is very 
    small. Hence, to assess the performance differences we first must rule out 
    small effects using Vargha and Delaney's A12 test, a test   recently endorsed by Arcuri and 
    Briand at ICSE'11~\cite{arcuri11}.

    To  apply Scott-Knott, we divided data into the phases $P_0$ where issues are introduced. Next, for each division, we separated all the issues that were removed at different subsequent issues $P_r \in \{P_1,P_2,..\}$.
    For each pair $P_0,P_r$, we build one treatment containing the  issue resolution times for     issues raised in $P_0$ and resolved in $P_r$. These treatments
    were then ranked by Scott-Knott.

@@

\multicolumn{7}{c}{~}  \\
1 &  Before   
      &   DesignInspect   & 10 &  14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   CodeReview      & 8 &  14 & 0.80 & \textcolor{black}{\rule{8mm}{2mm}}  \\
1 &   &   CodeInspect     & 10 &  16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   UnitTest        & 12 &   21 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &   &   IntTest         & 15 &   31 & 1.50 & \textcolor{black}{\rule{15mm}{2mm}}  \\
1 &   &   SysTest         & 11 &   22 & 1.10 & \textcolor{black}{\rule{12mm}{2mm}}  \\  
\hline\multicolumn{7}{c}{~}  \\
1 &  Planning     &   ReqtsReview     & 8 &  14 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &               &   DesignInspect   & 11 &  13 & 1.38 & \textcolor{black}{\rule{13mm}{2mm}} \\
2 &               &   UnitTest        & 24 &   25 & 3.00 & \textcolor{black}{\rule{30mm}{2mm}} \\
\hline\multicolumn{7}{c}{~}  \\
1 &  Reqts   &   ReqtsReview     & 13 &  20 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} \\
1 &          &   ReqtsInspect    & 12 &   18 & 0.92 & \textcolor{black}{\rule{9mm}{2mm}}  \\
1 &          &   DesignReview    & 10 &   14 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}} \\
1 &          &   DesignInspect   & 9 &   15 & 0.69 & \textcolor{black}{\rule{6mm}{2mm}}   \\
1 &          &   CodeInspect     & 13 &  24 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}} \\
1 &          &   UnitTest        & 10 &  17 & 0.77 & \textcolor{black}{\rule{7mm}{2mm}}  \\
1 &          &   IntTest         & 33 &   42 & 2.54 & \textcolor{black}{\rule{25mm}{2mm}}  \\
1 &          &   SysTest         & 24 &  108 & 1.85 & \textcolor{black}{\rule{18mm}{2mm}}  \\
\hline\multicolumn{7}{c}{~}  \\
1 & Design   &   DesignReview    & 11 &   16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   DesignInspect   & 8 &   12 & 0.73 & \textcolor{black}{\rule{7mm}{2mm}}  \\
1 &   &   CodeReview      & 10 &   18 & 0.91 & \textcolor{black}{\rule{9mm}{2mm}}  \\
1 &   &   CodeInspect     & 9 &  14 & 0.82 & \textcolor{black}{\rule{8mm}{2mm}}  \\
1 &   &   UnitTest        & 11 &   18 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   IntTest         & 17 &   31 & 1.55 & \textcolor{black}{\rule{15mm}{2mm}}  \\
1 &   &   SysTest         & 13 &   18 & 1.18 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &   &   AcceptTest      & 14 &   19 & 1.27 & \textcolor{black}{\rule{12mm}{2mm}}  \\
\hline\multicolumn{7}{c}{~}  \\
1 &  Code   &   CodeReview    & 10 &  16 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &    &   CodeInspect   & 10 &   15 & 1.00 & \textcolor{black}{\rule{10mm}{2mm}}  \\
1 &   &   UnitTest      & 12 &   20 & 1.20 & \textcolor{black}{\rule{12mm}{2mm}}  \\
1 &    &   IntTest       & 14 &   25 & 1.40 & \textcolor{black}{\rule{14mm}{2mm}}  \\
1 &   &   SysTest       & 13 &   20 & 1.30 & \textcolor{black}{\rule{13mm}{2mm}}  \\
1 &    &   AcceptTest    & 16 &   25 & 1.60 & \textcolor{black}{\rule{16mm}{2mm}}  \\

@@

\end{center}
\caption{Times to resolve issues seen in the SEI TSP data. For an explanation of this figure,
see \tion{171projects}.}
\label{fig:raw}
\end{figure}

@@

\textit{Reliability of measures}: The base measures in this study described in \tion{data-collection} are \textit{defects} recorded in TSP defect logs and \textit{time} reported in time tracking logs. The primary threats to the reliability of these measures are: that the definition of a defect varies between projects and that time is not reported accurately or consistently. The reliability of the time reporting is discussed in \tion{data_integrity}. Time is reported on a level of minutes. We do not have a precise assessment of the error margin for time reporting. Some developers are less precise with time or estimates. Nonetheless, we have applied several tests to verify that the data is accurate. First we compare entries from the defect and time logs to verify that defect log times-to-fix sum to less than the total time log effort in the phase. Second, time log time stamps must be consistent with both the the time stamps and phase for defect in the defect log. Third, we applied a Benford test on the leading digits from the time log  and defect log times to estimate the number data entries that do not result from a natural process (that is, guessed or estimated rather than measured values) ~\cite{shirai14}.  Based on these tests we believe that greater than 90\% of the time log data is recorded in real time. The fidelity and consistency of data will be subject of a future paper.
